Mechanism to control synchronization in coupled area-preserving maps

### 2018-08-29 Syncarea_control/paper.tex
paper updated

%
%   This file is part of the APS files in the REVTeX 4 distribution.
%   Version 4.0 of REVTeX, August 2001
%
%   Copyright (c) 2001 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.0
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%                                                               /Users/neelimamgupte/Desktop/iitm.eps                                                                                                                                                                                 
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}

% Some other (several out of many) possibilities
%\documentclass[preprint,aps]{revtex4-1}
%\documentclass[preprint,aps,draft]{revtex4}
%\documentclass[prb]{revtex4-1}% Physical Review B
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{amsmath,empheq}
\usepackage{hyperref}
\usepackage{natbib}
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\begin{document}
\title{A mechanism to control synchronization in coupled area-preserving maps}
\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}
\author{Swetamber Das}
\email{swetdas@pks.mpg.de}
\affiliation{Max Planck Institute for the Physics of Complex Systems \\ 
N\"othnitzer Stra\ss e 38, Dresden, Germany.}
%\author{Neelima Gupte}
%\email{gupte@physics.iitm.ac.in}
%\affiliation{Department of Physics, Indian Institute of Technology Madras, Chennai, 600036, India.}

\begin{abstract}
    A unidirectionally coupled area-preserving maps with mixed phase space may 
    show identical synchronization in the sticky neighborhood of the regular 
    islands. We use this fact to devise a couple of numerical procedures to 
    delay and expedite the process of synchronization in standard maps coupled 
    under  Pecora-Caroll scheme. The delay method is based on controlled 
    kicking of trajectories away from the regular islands for as long as 
    necessary. On the other hand, the method to expiate the process is
    achieved by a parameter perturbation technique which rapidly drives the 
    chaotic trajectories to synchronization traps in the sticky neighborhoods 
    of regular islands. We also point out of limitations of these methods. 
\pacs{05.45.+b}
\end{abstract}
\maketitle

@@

Synchronization in coupled chaotic systems, coupled map lattices and networks 
have been studied in great details over the years. The fact however remains 
that the ability to control synchronization based on the dynamics of the 
process has not been given much attention, until 
recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
processing of visual or  synchronization may be significant. For example, in 
neurosciences,  the speed of the visual and olfactory processing are 
interesting problems \cite{Thorpe1996,Uchida2003}. In this paper, we intend 
to develop a couple of 
mechanisms to control synchronization process to wit,  to increase and to 
decrease synchronization time based on the location of synchronization traps 
in the sticky neightborhood of regular islands in the phase space of the 
area-preserving system of the standard map. We demonstrate the mechanism at $K 
= 6$ where only two small regular islands exist.  The rest of the paper is 
organized in the following way: we explain the coupling scheme in 
sec.~\ref{sec:Pecora_Carroll}, the mechanism and demonstration of delay and 
advanced synchronization times in sec.~\ref{sec:delay} and 
sec.~\ref{sec:advanced} respectively, and the paper ends with conclusions in 
sec.~\ref{sec:conclusions} including discussion on limitations and 
implications.

@@

\begin{figure}[b]
\includegraphics[scale=.6]{Standard_Map}
\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
\end{figure}
Here the subscript $n$ denotes the discrete time and $K$ is the nonlinearity parameter. These equations typically describe the evolution of two canonical variables $P$ and $Q$ which correspond to the momentum and co-ordinate in the Poincar\'{e} section of a freely moving rotator with interleaved periodic kicks. This system represents the behavior of a variety of systems such as charged particle confinement in  mirror magnetic traps, particle dynamics in accelerator, comet dynamics in solar systems etc.  Two-dimensional  phase space plots  of the standard map for the parameter value $K = 6$ using $25$ initial conditions  are shown in Figs.~\ref{fig:Standard_map}.

@@

\begin{figure}[t]
\includegraphics[scale=.6]{Sync_exmaple}
\caption{\label{fig:sync_ex} \footnotesize The variation of $\Delta Q = Q^d_n-Q^r_n$ of the coupled system at $K = 6$ with iterations $n$. Synchronization is considered to be achieved when $\Delta Q < 10^{-5}$ and remain so for the rest of the iterations. The iteration step $n$ at which synchronization occurs is called synchronization time $\tau$. For this case, $\tau = 570 246.$ }
\end{figure}

@@

\section{Locating Synchronzation Traps using Master stability function}
\label{Master stability function}
A general drive-response system coupled unidirectionally may be described by the following set of equations

@@

Therefore, in the case of a general unidirectional coupling of two standard maps, we get

\begin{center}
\begin{empheq}[right=\empheqrbrace \mod 1]{align}
P^d_{n+1} = P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} = P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} = P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} = P^r_{n+1} + Q^r_n 
\label{equ:coupled}
\end{empheq}
\end{center}

@@

\begin{figure}[h]
\includegraphics[scale=0.55]{FTLE_plots}
\caption{\label{fig:FTLE}\footnotesize Fluctuations of time-5 FTLEs for three different set of initial conditions -- $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c). The synchronization times in each cases are 287090, 234521, and 164782 respectively. The time scale on the x-axis indicates iterates steps $t$ at which time-5 FTLEs have been computed. The drop window in (a), (b), and (c) shown in ellipses in red occur around synchronization times. Synchronization of chaotic trajectories in the coupled system under study typically occurs in one of such traps.}
\end{figure}

@@

\begin{figure}[h]
\includegraphics[scale=0.6]{Edge_Sync_angles}
%\includegraphics[scale=.4]{Standard_Map2.eps}\\
\caption{\label{fig:Edge_Sync_angles} \footnotesize Numerical detection of 
synchronization traps (a) Distribution of phase 
angles of the points on the edges of regular islands determined by the edge 
detection algorithm. (b) Distribution of phase angles for points where 
synchronization occurs. }
\end{figure}

@@

The proximity parameter $d_0 = 0.02$ defines the region in the neighbourhood of the numerically detected edges wherein synchronization traps exist. The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighbourhood followed by a slight deflection so that the trajectory resters at a point outside.  Our procedure depends upon how many times we deflect the trajectory  which will be referred to as step control. This means that if an $n-$step control is  employed then, the trajectory will be kicked away from the domain $n$-times during its first $n$ visits i.e. once per visit. The deflection is achieved by adding a randomly generated fraction below $0.1$ to the point visiting the domain, and therefore the maximum deflection area is roughly $1\%$ of the phase space.  This procedure applied on a given set of initial conditions of the drive and response maps, may result in four possibilities of synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay \textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) undesirable faster synchronization \textbf{(U)}. Clearly, the only first possibility is the aim of the mechanism; the last one constitute a hazard while the other two are trivial. 
\begin{figure}[h]
\includegraphics[scale=0.6]{S_U_Percent}
\caption{\label{fig:Control_success} \footnotesize. The curves indicate the 
percentages of (S) successful delays shown as the blue curve,a and (U) 
undesirable fast synchronization showns as the red curve. }
\end{figure}

@@

\begin{figure}[t]
\includegraphics[scale=0.6]{Sync_time_dist.eps}
\caption{\label{fig:Sync_time_dist} \footnotesize. Distribution of 
synchronization time without n-step control (the blue curve) and with n-step 
control (the red curve). The control delays synchronization times. }
\end{figure}

@@

 \begin{figure}[t]
\includegraphics[scale=0.6]{Strength_con.eps}
\caption{\label{fig:Strength_con} \footnotesize. Variation of $\beta$ 
(maximum, minimum and average) which measures the strength of successful 
delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$. $\beta_max$ indicates the 
maximum delay time obtained at a give n-step control while $\beta_{mean}$ 
stands for the mean value around 2. This means that on an average, 
synchronization times re delayed by 100 times. The minimum value of $\beta$ 
remains 1 for the cases when no delay was obtained obtained. }
\end{figure}

  The extent of delay $\beta$ may be defined as the ratio of successfully 
  delayed synchronization times $\tau_c$ to synchronization times $\tau_0$ 
  without control, i.e. $\beta =  \ln(\frac{\tau_d}{\tau_0}$).  Therefore, for 
  successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$. 
  The plot in Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. 
  $\beta_{min}$, $\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean 
  and maximum $\beta$ for considered $n-$step controls. Clearly, we are able 
  to attain $\beta_{max}$ values to be in the range $5$ to $14$ while mean 
  values remain $\sim2$. Therefore, we are successful in achieving 
  significantly large delays in synchronization times.

@@

 \section{Advanced Syncronization Times}
 \label{sec:advanced}
 We now explore the possibility to decrease synchronization times i.e. to expedite the process of synchronization. Once again, we will make use of the fact that synchronization traps typically exist in the sticky neighbourhoods of the regular islands.  The procedure is based on the parameter perturbation technique to generate coherent structures in the phase space of an area-preserving map. However, the technique may also be used to push the chaotic bulk on the coherent structure (regular region) already present in the phase space and therefore, to make it disappear completely.  We describe the procedure briefly.

@@

\begin{minipage}[t]{0.55\textwidth}
%\raggedleft
\centering
\[ J = \left( \begin{array}{cc}
1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right)\]  
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\raggedright

\vspace{0.65cm}
$ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\
\end{minipage}

\vspace{0.65cm}
The determinant of this matrix $J$ remains 1. The phase space thus obtained for for $K = 6$ , $\delta = 0.3$ and $\epsilon = 2$ for the fixed point $(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:coherent}.   We have performed computation for the other fixed point and with other modulo operations as well and results do not differ significantly. Our choice here of modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the others in terms of actual numerical values of synchronization times and the display of locations of the points where synchronization occurs.

For the computations, we take 50 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map.  The procedure has been applied to both drive and response map in accordance with our coupling  scheme described above. We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.3$ which corresponds to about $9\%$ of the phase space area. We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} wherein synchronization occurs in the neighbourhood of the main regular island and one of its satellite islands.  The sign `+' in red indicate the point of synchronization. 
 \begin{figure}[h]
\includegraphics[scale=0.6]{Fast_sync_location.eps}
\caption{\label{fig:location} \footnotesize.The phase space of the drive map 
at K = 6 where parameter perturbation is applied. Synchronization occors at 
the location indicate by the red 'plus' signs.}
\end{figure}

 \begin{figure}[h]
\begin{tabular}{cc}
\includegraphics[scale=0.45]{Normal_Sync.eps}&
\includegraphics[scale=0.45]{Fast_sync_dist.eps}\\
(a) & (b)
\end{tabular}{}
\caption{\label{fig:sync_time_dist} \footnotesize The phase space of the standard map for 25 random initial conditions from the uniform distribution on $[0,2\pi]$ plotted for 4500 iterations; 500 iterates are discarded as transients. The values are normalized i.e. $\frac{P}{2\pi} $ and $\frac{Q}{2\pi}$ are plotted on the $y$ and $x$ axis for the parameter values (a)  K = 1.5, and (b) K = 6.0.}
\end{figure}

The probability distribution of synchronization times without parameter perturbation shows a long tail which shows a power law scaling. It is shown using cumulative probability distribution on a log-log scale in Fig.~\ref{fig:sync_time_dist}(a) with slope $\sim 0.744$. Upon application of perturbation, synchronization times become considerable shorter and the probability distribution cross over to exponential. Fig.~\ref{fig:sync_time_dist}(b) shows the resulting exponential distribution for the same set of initial conditions as shown for in Fig.~\ref{fig:sync_time_dist}(a) , with an estimated mean $\sim500$. Comparing the $x-$scales of the plots, one can see the faster synchronization has been successfully achieved.

We again define the strength of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$. For advanced synchronization, $\beta < 0$ indicating that synchronization times upon parameter perturbation is smaller than in normal course i.e. $\tau_c < \tau$. The probability distribution of $\beta$ in shown in Fig.~\ref{fig:beta_dist}, which may be fit with a gaussian of the following form:
\begin{equation}
f(\beta) =  a\exp(-\frac{(\beta-\mu)^2}{2c^2})
\end{equation}

We estimate the mean value $\mu \sim - 6.648$ and standard deviation $\sigma \sim1.542$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are $4\times10^2$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^4$ times smaller.  Therefore, the synchronization times we obtained upon parameter perturbation is significantly smaller than the normal synchronization times.

The parameter perturbation technique is an efficient way to decrease synchronization times with high accuracy.  The value of the perturbation imposed should be chosen suitably as a very small may not have any influence on the synchronization process. The procedure also depends on the size and location of the patch wherein the control is applied. We are yet to see if the other technique of generating coherent structures also influence synchronization in these unidirectionally coupled systems.

 \begin{figure}[h]
\includegraphics[scale=0.6]{Fast_strength_dist.eps}
\caption{\label{fig:location} \footnotesize.}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
To summarise, we have developed a numerical procedure to delay synchronization in two identical standard map coupled under the drive-response configuration. The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighbourhoods of the regular islands in the phase space. We have shown that the delay is obtained by kicking the trajectories slightly away from the domain temporarily. The efficiency of the procedure depends upon the number of steps employed in control. By applying the procedure to the system on merely $1\%$ of the phase space area (at $K = 6$), we have successfully increased synchronization times. The maximum number of control steps that are used is $0.01\%$ of the total of number of iterations used throughout the computation.  Furthermore, the procedure gives variety of delays. The limitations of the procedure are, one, the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization and two, a predetermined synchronization time in not obtainable.

%\begin{thebibliography}
\bibliographystyle{unsrt}
\bibliography{Syncref}

%\end{thebibliography}
\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% **IMPORTANT** remarks ahead
% - linelength <= 80 (in the text, unless unavoidable)
% - no TABs
% - no trailing spaces
% - Use macros for more complicated variables
%   (especially, if the notation might be changed later...)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[reprint,superscriptaddress,amsmath,amssymb,aps,pre]{revtex4-1}

\usepackage{graphicx}% Include figure files
\graphicspath{plots}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{bbold}
\usepackage{soul}

\usepackage{empheq}
\setcounter{topnumber}{1}

\usepackage{flafter}     % Bilder immer nach \figure Befehl

% http://tex.stackexchange.com/questions/62010/can-i-access-system-environment-variables-from-latex-for-instance-home
\usepackage{catchfile}
\newcommand{\getenv}[2][]{%
\CatchFileEdef{\temp}{"|kpsewhich --var-value #2"}{\endlinechar=-1}%
\if\relax\detokenize{#1}\relax\temp\else\let#1\temp\fi}

%\getenv[\HOME]{HOME}

% use okular to view this file and acroread will be called on links
%\newcommand{\articlestoragepath}{/home/scratch/abaecker/ArticleStorage/}
\getenv[\articlestoragepath]{ARTICLE_STORAGE}
% Conditionally import pdf_article_link.sty to ensure that it works
% even without this file.
\IfFileExists{pdf_article_link.sty}{\usepackage{pdf_article_link}}

\input defs.tex
\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADD THE FOLLOWING COUPLE LINES INTO YOUR PREAMBLE
\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

\makeatletter
\let\Hy@backout\@gobble
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Mechanism to control synchronization in coupled area-preserving maps}

\author{Swetamber Das}
\affiliation{Max-Planck-Institut f\"ur Physik komplexer Systeme, N\"othnitzer
    Stra\ss{}e 38, 01187 Dresden, Germany}

\date{\today}

\begin{abstract}
A unidirectionally coupled area-preserving maps with mixed phase space may identical synchronization in the sticky neighborhood of the regular islands. We use this fact to devise numerical procedures to delay and advance the process of synchronization. The delay method is based on controlled kicking of the trajectories away from the regular islands for as long as required. The method to expedite the process is achieved by a parameter perturbation technique which rapidly drives the chaotic trajectories in the neighborhood of regular islands where synchronization occurs. We also point out the limitations of these methods.   
\end{abstract}

\pacs{PACS here}

\maketitle

@@

Synchronization in coupled chaotic systems, coupled map lattices and networks 
have been studied in great details over the years. Coupled Hamiltonian systems are also known to show synchronization such as measure synchronization~\cite{Hampton1999,Wang2003,Vincent2005,Gupta2017} and identical synchronization~\cite{Mahata2016,Das2017} The fact however remains 
that the ability to control synchronization based on the dynamics of the 
process has not been given much attention, until 
recently~\cite{Grabow2011,Wang2016}. In realistic systems, the speed of 
processing of visual or  synchronization may be significant. For example, in 
neurosciences,  the speed of the visual and olfactory processing are 
interesting problems \cite{Thorpe1996,Uchida2003}. In this paper, we intend 
to develop a couple of 
mechanisms to control synchronization process to wit,  to increase and to 
decrease synchronization time based on the location of synchronization traps 
in the sticky neightborhood of regular islands in the phase space of the 
area-preserving system of the standard map. We demonstrate the mechanism at $K 
= 6$ where only two small regular islands exist. It is important to note that role of stickiness in Hamiltonian systmes for synchronaization was already predicted by Zaslavsky~\cite{Zaslavsky2002b} early in the last decade.\\

The rest of the paper is 
organized in the following way: we explain the coupling scheme in 
sec.~\ref{sec:Pecora_Carroll}, the mechanism and demonstration of delay and 
advanced synchronization times in sec.~\ref{sec:delay} and 
sec.~\ref{sec:advanced} respectively, and the paper ends with conclusions in 
sec.~\ref{sec:conclusions} including discussion on limitations and 
implications.

@@

\begin{figure}[b]
	\includegraphics[scale=.45]{Standard_Map}
	\caption{\label{fig:Standard_map} \footnotesize The phase space of the standard map for  random initial conditions from the uniform distribution on $[0,1]$,  at the parameter value K = 6.0.}
\end{figure}
Here the subscript $n$ denotes the discrete time and $K$ is the nonlinearity parameter. These equations typically describe the evolution of two canonical variables $P$ and $Q$ which correspond to the momentum and co-ordinate in the Poincar\'{e} section of a freely moving rotator with interleaved periodic kicks. This system represents the behavior of a variety of systems such as charged particle confinement in  mirror magnetic traps, particle dynamics in accelerator, comet dynamics in solar systems etc.  Two-dimensional  phase space plots  of the standard map for the parameter value $K = 6$ using $25$ initial conditions  are shown in Figs.~\ref{fig:Standard_map}.

@@

\begin{figure}[t]
	\includegraphics[scale=.4]{Sync_exmaple}
	\caption{\label{fig:sync_ex} \footnotesize The variation of $\Delta Q = Q^d_n-Q^r_n$ of the coupled system at $K = 6$ with iterations $n$. Synchronization is considered to be achieved when $\Delta Q < 10^{-5}$ and remain so for the rest of the iterations. The iteration step $n$ at which synchronization occurs is called synchronization time $\tau$. For this case, $\tau = 570 246.$ }
\end{figure}

@@

\section{Locating Synchronization Traps using Master stability function}
\label{Master stability function}
A general drive-response system coupled unidirectionally may be described by the following set of equations

@@

Therefore, in the case of a general unidirectional coupling of two standard maps, we get
\begin{align}
P^d_{n+1} &= P^d_n + \frac{K}{2\pi}\sin(2\pi Q^d_n) \nonumber\\
Q^d_{n+1} &= P^d_{n+1} + Q^d_n \nonumber\\
P^r_{n+1} &= P^r_n + \frac{K}{2\pi}\sin (2\pi Q^r_n) + \alpha(P^d_n-P^r_n) \nonumber \\
Q^r_{n+1} &= P^r_{n+1} + Q^r_n 
\label{equ:coupled}
\end{align}

@@

\begin{figure*}[t!]
	\includegraphics[scale=0.41]{FTLE_plots}
	\caption{\label{fig:FTLE}\footnotesize Fluctuations of time-5 FTLEs for three different set of initial conditions -- $(P^d_0,Q^d_0,Q^r_0) = (0.466,0.098,0.141)$ in (a),  $(0.009,0.026,0.129)$ in (b), and $(0.553,0.554,0.396)$ in (c). The synchronization times in each cases are 287090, 234521, and 164782 respectively. The time scale on the x-axis indicates iterates steps $t$ at which time-5 FTLEs have been computed. The drop window in (a), (b), and (c) shown in ellipses in red occur around synchronization times. Synchronization of chaotic trajectories in the coupled system under study typically occurs in one of such traps.}
\end{figure*}

@@

\begin{figure}[h]
	\includegraphics[scale=0.45]{Edge_Sync_angles}
	%\includegraphics[scale=.4]{Standard_Map2.eps}\\
	\caption{\label{fig:Edge_Sync_angles} \footnotesize Numerical detection of 
		synchronization traps (a) Distribution of phase 
		angles of the points on the edges of regular islands determined by the edge 
		detection algorithm. (b) Distribution of phase angles for points where 
		synchronization occurs. }
\end{figure}

@@

The proximity parameter $d_0 = 0.02$ defines the region in the neighbourhood of the numerically detected edges wherein synchronization traps exist. The basic idea to control synchronization time is to identify the step at which the chaotic trajectory visits the numerically determined sticky neighborhood followed by a slight deflection so that the trajectory restarts at a point outside.  Our procedure depends upon how many times we deflect the trajectory  which will be referred to as step control. This means that if an $n-$step control is  employed then, the trajectory will be kicked away from the domain $n$-times during its first $n$ visits i.e. once per visit. The deflection is achieved by adding a randomly generated fraction below $0.1$ to the point visiting the domain, and therefore the maximum deflection area is roughly $1\%$ of the phase space.  This procedure applied on a given set of initial conditions of the drive and response maps, may result in four possibilities of synchronization time -- (1) successful delay  \textbf{(S)}, (2) no delay \textbf{(N)}, (3) failed to achieve synchronization \textbf{(F)}, and (4) undesirable faster synchronization \textbf{(U)}. Clearly, the only first possibility is the aim of the mechanism; the last one constitute a hazard while the other two are trivial. 
\begin{figure}[h]
	\includegraphics[scale=0.45]{S_U_Percent}
	\caption{\label{fig:Control_success}\footnotesize The curves indicate the 
		percentages of (S) successful delays shown as the blue curve,a and (U) 
		undesirable fast synchronization shown as the red curve. }
\end{figure}

@@

\begin{figure}[t]
	\includegraphics[scale=0.45]{Sync_time_dist.eps}
	\caption{\label{fig:Sync_time_dist}\footnotesize Distribution of 
		synchronization time without n-step control (the blue curve) and with n-step 
		control (the red curve). The control delays synchronization times. }
\end{figure}

@@

\begin{figure}[t]
	\includegraphics[scale=0.45]{Strength_con.eps}
	\caption{\label{fig:Strength_con} \footnotesize. Variation of $\beta$ 
		(maximum, minimum and average) which measures the strength of successful 
		delays i.e. $\beta = \ln(\frac{\tau_d}{\tau_0}$. $\beta_max$ indicates the 
		maximum delay time obtained at a give n-step control while $\beta_{mean}$ 
		stands for the mean value around 2. This means that on an average, 
		synchronization times re delayed by 100 times. The minimum value of $\beta$ 
		remains 1 for the cases when no delay was obtained obtained. }
\end{figure}

The degree of delay $\beta$ may be defined as the ratio of synchronization times without numerical control, $\tau_c$ to synchronization times $\tau_0$ without control, i.e. $\beta =  \ln(\frac{\tau_d}{\tau_0}$).  Therefore, for successfully delayed synchronization $\beta>0$ indicating $\tau_c > \tau$. 
The plot in Fig.~\ref{fig:Strength_con} shows the variation of $\beta$ i.e. 
$\beta_{min}$, $\beta_{mean}$, and $\beta_{max}$  indicating minimum, mean 
and maximum $\beta$ for considered $n-$step controls. Clearly, we are able 
to attain $\beta_{max}$ values to be in the range $5$ to $14$ while mean 
values remain $\sim2$. Therefore, we are successful in achieving 
significantly large delays in synchronization times.

@@

\section{Advanced Synchronization Times}
\label{sec:advanced}
Based on the now explore the possibility to decrease synchronization times i.e. to expedite the process of synchronization. Once again, we will make use of the fact that synchronization traps typically exist in the sticky neighbourhoods of the regular islands.  The procedure is based on the parameter perturbation technique to generate coherent structures in the phase space of an area-preserving map. However, the technique may also be used to push the chaotic bulk on the coherent structure (regular region) already present in the phase space and therefore, to make it disappear completely.  We describe the procedure briefly.

@@

\begin{minipage}[t]{0.45\textwidth}
	%\raggedleft
	\centering
	\[ J = \left( \begin{array}{cc}
	1 & 1 + (K-\epsilon)\cos(2\pi Q_n)   \\
	1 & (K-\epsilon)\cos(2\pi Q_n) \end{array} \right)\] 
\end{minipage}

\vspace{0.4cm}
	where, $ \epsilon \neq 0$ if $|P_f - P_n| < \delta$, $|Q_f - Q_n| < \delta$\\ $\epsilon = 0$, otherwise.\\

\begin{figure*}[t]
	\includegraphics[scale=0.4]{Fast_sync_location.eps}
	\caption{\label{fig:location}\footnotesize The phase space of the drive map at K = 6 for a couple of initial conditions where parameter perturbation is applied. Synchronization occurs at the location indicate by the red 'plus' signs.}
\end{figure*}

The determinant of this matrix $J$ remains 1. The phase space thus obtained for for $K = 6$ , $\delta = 0.3$ and $\epsilon = 2$ for the fixed point $(P_f,Q_f)=(0.0,0.0)$ is shown in Fig.~\ref{fig:coherent}.   We have performed computation for the other fixed point and with other modulo operations as well and results do not differ significantly. Our choice here of modulo operation and the fixed point $(P_f,Q_f)$ has some advantage over the others in terms of actual numerical values of synchronization times and the display of locations of the points where synchronization occurs.

For the computations, we take 5 000 initial conditions chosen randomly from the chaotic region of the unperturbed standard map i.e. $P_0 \in \{-0.4,0.4\}$ and $Q_0 \in \{0.3,0.7\}$.   The procedure has been applied to both drive and response map in accordance with our coupling  scheme described above. We choose a patch around $(P_f,Q_f)$  of length $\delta= 0.4$ which corresponds to about $16\%$ of the phase space area. We successfully obtain advance synchronization in more that $99.5\%$ of the cases among which a  couple of instances  has been shown in Fig.~\ref{fig:location} -- synchronization occurs in the neighborhood of the main regular island and one of its satellite islands.  The sign `+' in red indicate the point of synchronization.

We again define the degree of controlled synchronization as $\beta = \ln \frac{\tau_c}{\tau}$. For advanced synchronization, $\beta < 0$ indicating that synchronization times upon parameter perturbation is smaller than in normal course i.e. $\tau_c < \tau$. The probability distribution of $\beta$ in shown in Fig.~\ref{fig:beta_dist}, which may be fit with a gaussian of the following form:
\begin{equation}
f(\beta) =  a\exp\Big(-\frac{(\beta-\mu)^2}{2c^2}\Big)
\end{equation}

We estimate the mean value $\mu \sim -7.46$ and standard deviation $\sigma \sim1.67$ which corresponds to the fact that, on an average, advanced synchronization times thus obtained are about $10^3$ times smaller than the normal synchronization times with $95.5\%$ values are between $10$ to $10^5$ times smaller.  Therefore, the synchronization times we obtained upon parameter perturbation is significantly smaller than the normal synchronization times.

The parameter perturbation technique is an efficient way to decrease synchronization times with high accuracy.  The value of the perturbation imposed should be chosen suitably as a very small may not have any influence on the synchronization process. The procedure also depends on the size and location of the patch wherein the control is applied. We are yet to see if the other technique of generating coherent structures also influence synchronization in these unidirectionally coupled systems. 
\begin{figure}[b]
	\includegraphics[scale=0.45]{sync_time_dist_beta.pdf}
	\caption{\label{fig:beta_dist}\footnotesize Distribution of degree of synchronization, $P(\beta)$.}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
To summarize, we have developed numerical procedures to control synchronization in two identical standard map coupled under the drive-response configuration. The procedure is based on the fact that synchronization in the system typically occurs in the sticky neighborhoods of the regular islands in the phase space. We have shown that a delay in synchronization can be obtained by kicking the trajectories slightly away from the domain temporarily. The efficiency of the procedure depends upon the number of steps employed in control. By applying the procedure to the system on merely $1\%$ of the phase space area (at $K = 6$), we have successfully increased synchronization times. The maximum number of control steps that are used is $0.01\%$ of the total of number of iterations used throughout the computation.  Furthermore, the procedure gives variety of delays. The limitations of the procedure are, one, the deflection may insert the drive trajectory in the regular islands which would lead to undesirable faster synchronization and two, a predetermined synchronization time in not obtainable. Furthermore, we have demonstrated a simple way to fasten synchronization by the parameter perturbation technique which rapidly drives the chaotic trajectories in the sticky neighborhood of regular islands. The distribution of degree of synchronization times indicates the efficiency of the procedure. A possible future direction is to investigate the role of stickiness in measure synchronization in coupled Hamiltonian systems.

\begin{acknowledgments}
I would like to thank Neelima Gupte and Arnd B\"acker for their suggestions to improve the manuscript. 
\end{acknowledgments}

\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Some stuff for the appendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\bibliographystyle{cpg_unsrt_title}
%\bibliographystyle{cpg_unsrt_key_title}
%\bibliographystyle{cpg_unsrt_title2}
%\bibliographystyle{cpg_unsrt_key_title2}
%\bibliographystyle{unsrt}
\bibliographystyle{cpg_unsrt_title_for_phys_rev}
%\bibliographystyle{cpg_unsrt_key_title2} % to have labels + links

% See ~/Devel/Repositories/Bibtex/notes.txt
% if you need more
\bibliography{Syncref}

\end{document}
